Lawful Access: Arguments and Alternatives

### 2021-04-06 articles/tots-ieee/Article.tex
modifications to make paper stand-alone, including pulling in paragraphs from thesis intro

\title{Confronting Wicked Crypto}

@@

% NOTE: The first paragraph of the abstract was copied. The second I tweaked.

Public debate has resumed on the topic of lawful access (LA) to encrypted data, which refers to alternative means of
decryption intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at
the end of the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
technology is a wicked problem that has and will continue to evade a permanent solution. As policymakers consider next
steps, it is necessary that the technical community remain engaged. Although any LA framework would increase risk, the
magnitude of that increase varies greatly with the quality of the technical and regulatory approach. Furthermore, if one
considers hard-line legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed
LA may reduce risk overall.

@@

This chapter will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the
debate, but there are two important shortcomings of the maps presented here. First, argument maps should be used simply
as tools in the collaborative pursuit of the problem definition and solution. Maps assembled from research are not a
replacement for debate arising from live discussion. Second, \acp{argmap} may deceptively portray the strength of an
argument. Argument nodes do not indicate their strength or validity. If ten unsound arguments were presented against one
indisputable argument, the side with more arguments would, regardless of the arguments' integrity, appear stronger. With
those qualifications in mind, let us begin by examining the factors at the center of the conflict.

@@

Information security is the remaining central issue in the debate. As described in \mysec{sec-intro}, \ac{cryptography}
plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought to compromise the
cryptographic foundations of encryption; however, this is seen as too risky today. \mysec{sec-intro} introduced several
alternative technical approaches to \ac{EA}, but experts still emphasize that current systems cannot securely provide
the level of access \acl{LE} asks for \cite{abelson_2015} \cite{abelson_risks_1997}. Security was an afterthought in
early computing and networking designs. The field of cybersecurity is still equal parts art and science; mandated
\ac{EA} would put a heavy burden on a field still finding its legs.

% TODO: fix \mysec{sec-intro} reference in above paragraph
% TODO: fix \mysec{sec-intro} reference in above paragraph (there are two instances)

@@

\section{Going Dark vs. The Golden Age}

@@

Still, law enforcement's inability to access the tools of \ac{masssurv} cannot be taken for granted. It will likely
change---they already have access to mobile devices. A more conservative criticism of the ``Golden Age'' argument rests
in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement while
thwarting \ac{masssurv} subvert most of the ``Golden Age'' argument. \ac{EA} proposals with these aims will be explored
in \mysec{sec-intro}.

% TODO: fix \mysec{sec-intro} reference in above paragraph

@@

The last four maps framed the debate, presented leading arguments, and identified fallacious arguments. The next map
explores potential solutions. \myfig{fig-arg-measures} divides potential solutions into the categories of current
capabilities, legal measures, and \ac{EA}. Current capabilities can be implemented today, and include maintenance of the
status quo and increasing investment in current programs. Legal measures require a change in the text or application of
the law, but do not require fundamentally new technical capabilities. These include compelling passwords, which requires
legal clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. (These
non-\ac{EA} approaches were first introduced in \mysec{sec-intro}.) In this section, \ac{EA} is compared to its
alternatives. \ac{EA} variations are compared in the next map.

% TODO: fix \mysec{sec-intro} reference in above paragraph

It should be noted that the \ac{argmap} format works best when it displays the debate surrounding just one conclusion or
proposal. This allows adequate room for elaboration on points and counterpoints. When a map includes multiple
conclusions or proposals, it becomes a tangled web of connections, leaving no room for additional detail. For the sake
of brevity, this thesis compares \ac{EA}'s alternatives and variations in single maps. For a thorough discussion of both
\ac{EA}'s alternatives and variations, see the National Academies of Sciences, Engineering, and Medicine's 2018 report
titled \ii{Decrypting the Encryption Debate: A Framework for Decision Makers} \cite{committee_decrypting_2018}.

@@

Another \ac{DIM} \ac{EA} proposal involves adding authorities as a ``ghost'' participant in a conversation, analogous to
a silent listener on a conference call. This was notably suggested by UK \ac{GCHQ} officials Ian Levy and Crispin
Robinson \cite{levy_robinson_2018}. Their writeup is based on the same foundational principles as this thesis:

\begin{displayquote}
In any discussion of cyber security, details matter.

Unfortunately, it's the details that are missing from the discussion around lawful access to commodity end-to-end
encrypted services and devices (often called the ``going dark'' problem). Without details, the problem is debated as a
purely academic abstraction concerning security, liberty, and the role of government.

There is a better way that doesn’t involve, on one side, various governments, and on the other side lawyers,
philosophers, and vendors' PR departments continuing to shout at each other. If we can get all parties to look at some
actual detail, some practices and proposals---without asking anyone to compromise on things they fundamentally believe
in---we might get somewhere.
\cite{levy_robinson_2018}
\end{displayquote}

Although this proposal was raised with good intentions, it is nevertheless problematic. Proponents of the ghost user
proposal consider the uncompromised use of strong encryption as the proposal's great strength. Although this is a good
property, compromising a messaging app's authorization protocol is as dangerous as compromising its encryption
\cite{callas_1_2019}. Fundamentally, this proposal suffers from the same weakness as trusted-party key escrow---it
introduces an authorization vulnerability in the messaging service provider's platform, resulting in dangerous
centralized access capability \cite{schneier_ghost_2019}. It has a different form, but same nature as rejected key
escrow proposals.

@@

Device encryption, particularly in mobile phones, may actually pose the smallest impediment to investigation of any
encryption technology. This fact undercuts the argument for device key escrow. The ``Mass Extraction'' report clearly
describes the current state of affairs: device manufacturers are losing the exploit arms race, private companies are
commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory strategy is
leading to pervasive data collection with little oversight \cite{koepke_2020}. This state of affairs reinforces the
lessons of \mysec{sec-intro}---inaction \ii{is} action when facing \acp{wicked-prob}, and in this case, inaction has led
to an undesirable outcome. This situation does not negate the value of research into device-oriented \ac{EA}. This is
because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
accessing devices again, this issue will immediately resurface. This is likely why law enforcement continues to push for
\ac{EA} despite present circumstances. Additionally, a formalized process may reduce government abuse, which threatens
privacy and civil liberties more systematically than criminal abuse.

% TODO: fix \mysec{sec-intro} reference in above paragraph

@@

The next chapter focuses on one specific device key escrow proposal. It defines a threat model and analyzes Stefan
Savage's 2018 proposal, ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
\cite{savage_lawful_2018}.

>>>>

\title{Lawful Access: Arguments and Alternatives}

@@

% NOTE: The first paragraph of the abstract was copied. The second I tweaked.
Public debate has resumed on the topic of lawful access (LA) to encrypted data, which refers to alternative means of
decryption intended for law enforcement use. The resumption of this debate is not a renege on a resolute promise made at
the end of the 1990s ``crypto war''; rather, it represents a valid reassessment of optimal policy in light of changing
circumstances. The imbalance between privacy, access, and security in the context of constantly changing society and
technology is a wicked problem that has and will continue to evade a permanent solution. As policymakers consider next
steps, it is necessary that the technical community remain engaged. Although any LA framework would increase risk, the
magnitude of that increase varies greatly with the quality of the technical and regulatory approach. Furthermore, if one
considers hard-line legislative action and malicious abuse of cryptosystems as part of the threat model, well-designed
LA may reduce risk overall.

@@

In encryption policy, \ac{EA} is alternative means of
decryption intended for law enforcement use. Characterized by a dynamic technological environment, disagreement about
underlying values, and resistance to a clear solution, the debate on \ac{EA} will not go away.

The conflict at the heart of the encryption and \ac{EA} debate is this: the same cryptographic and design principles
that underlie nearly all digital security also enable an unprecedented degree of individual privacy. Encryption is a
foundational tool to the integrity and confidentiality of all connected computing systems. Its increasing ubiquity in
communications and storage provides clear benefits. In a world where information security is frightfully poor yet
increasingly important, the necessity of strong encryption cannot be understated. However, the privacy afforded by
certain encryption technologies hampers law enforcement investigations and hides wrongdoing \cite{cox_2020}
\cite{keller_internet_2019}. In a society that cares about bringing wrongdoers to justice, this risk should not be
ignored.

The conversation regarding encryption has reached a stalemate. Governments and law enforcement agencies frequently cite
the need to access encrypted data to perform their duties \cite{ministerial_2018} \cite{intl_2020} \cite{comey_2014}.
Human rights groups and technical leaders counter that a weakened encryption environment would fatally compromise
privacy and security as we know it \cite{abelson_2015} \cite{eightythree_2017} \cite{ruiz_there_2018}. As the debate
continues, many on both sides have become entrenched in their positions and have engaged in disingenuous attacks against
their opponents.

In the long term, increased regulation of the tech industry is inevitable and regulation of encryption is possible.
Despite the benefits of strong encryption, regulatory interest in the subject has not subdued. The form that regulation
will take in the field of encryption depends on the good faith efforts made to equitably balance the benefits and risks
involved in deploying a cryptographic system. If the technical community acts, it can lead and shape legislation rather
than be subjected to it.

Most importantly, in the pursuit of data privacy, regulatory action is part of the threat model. If the technical
community fails to act but regulators move forward, everyone may become subject to technically misguided, harmful laws.
Bad policy is just as much a threat as weak passwords. Due to this threat, as long as lawmakers continue pursuing
\ac{EA} regulation, it is the responsibility of the technical community to engage in discussion and respond to the
presented arguments.

This paper will analyze the \ac{encryption} debate using \acp{argmap}. The maps facilitate comprehension of the
debate, although two important shortcomings of the maps presented here must be acknowledged. First, argument maps should be used simply
as tools in the collaborative pursuit of the problem definition and solution. Maps assembled from research are not a
replacement for debate arising from live discussion. Second, \acp{argmap} may deceptively portray the strength of an
argument. Argument nodes do not indicate their strength or validity. If ten unsound arguments were presented against one
indisputable argument, the side with more arguments would, regardless of the arguments' integrity, appear stronger. With
those qualifications in mind, let us begin by examining the factors at the center of the conflict.

@@

Information security is the remaining central issue in the debate. \ac{cryptography}
plays a foundational role in nearly every aspect of security. Past \ac{EA} regulation efforts sought to compromise the
cryptographic foundations of encryption; however, this is seen as too risky today. Although there are several alternative technical approaches to \ac{EA}, experts still emphasize that current systems cannot securely provide
the level of access \acl{LE} asks for \cite{abelson_2015} \cite{abelson_risks_1997}. Security was an afterthought in
early computing and networking designs. The field of cybersecurity is still equal parts art and science; mandated
\ac{EA} would put a heavy burden on a field still finding its legs.

@@

\section{The privacy vs safety debate}

@@

Still, law enforcement's inability to access the tools of \ac{masssurv} cannot be taken for granted. It will likely
change---they already have access to mobile devices. A more conservative criticism of the ``Golden Age'' argument rests
in the fact that \ac{EA} without \ac{masssurv} is conceivable. \ac{EA} implementations that enable law enforcement while
thwarting \ac{masssurv} subvert most of the ``Golden Age'' argument.

@@

The last four maps framed the debate, presented leading arguments, and identified fallacious arguments. The next map
explores potential solutions. \myfig{fig-arg-measures} divides potential solutions into the categories of current
capabilities, legal measures, and \ac{EA}. Current capabilities can be implemented today, and include maintenance of the
status quo and increasing investment in current programs. Legal measures require a change in the text or application of
the law, but do not require fundamentally new technical capabilities. These include compelling passwords, which requires
legal clarification, and sanctioning lawful hacking, which requires a strategic pivot and an oversight framework. In this section, \ac{EA} is compared to its
alternatives. \ac{EA} variations are compared in the next map.

It should be noted that the \ac{argmap} format works best when it displays the debate surrounding just one conclusion or
proposal. This allows adequate room for elaboration on points and counterpoints. When a map includes multiple
conclusions or proposals, it becomes a tangled web of connections, leaving no room for additional detail. For the sake
of brevity, this paper compares \ac{EA}'s alternatives and variations in single maps. For a thorough discussion of both
\ac{EA}'s alternatives and variations, see the National Academies of Sciences, Engineering, and Medicine's 2018 report
titled \ii{Decrypting the Encryption Debate: A Framework for Decision Makers} \cite{committee_decrypting_2018}.

@@

Another \ac{DIM} \ac{EA} proposal involves adding authorities as a ``ghost'' participant in a conversation, analogous to
a silent listener on a conference call. This was notably suggested by UK \ac{GCHQ} officials Ian Levy and Crispin
Robinson \cite{levy_robinson_2018}. Although this proposal was raised with good intentions, it is nevertheless problematic. Proponents of the ghost user
proposal consider the uncompromised use of strong encryption as the proposal's great strength. Although this is a good
property, compromising a messaging app's authorization protocol is as dangerous as compromising its encryption
\cite{callas_1_2019}. Fundamentally, this proposal suffers from the same weakness as trusted-party key escrow---it
introduces an authorization vulnerability in the messaging service provider's platform, resulting in dangerous
centralized access capability \cite{schneier_ghost_2019}. It has a different form, but same nature as rejected key
escrow proposals.

@@

Device encryption, particularly in mobile phones, may actually pose the smallest impediment to investigation of any
encryption technology. This fact undercuts the argument for device key escrow. The ``Mass Extraction'' report clearly
describes the current state of affairs: device manufacturers are losing the exploit arms race, private companies are
commodifying the \ac{lawful-hacking} approach for agencies large and small, and the lack of regulatory strategy is
leading to pervasive data collection with little oversight \cite{koepke_2020}. 
This situation does not negate the value of research into device-oriented \ac{EA}. This is
because the advantage hackers hold over device manufacturers is temporary; when law enforcement faces difficulty
accessing devices again, this issue will immediately resurface. This is likely why law enforcement continues to push for
\ac{EA} despite present circumstances. Additionally, a formalized process may reduce government abuse, which threatens
privacy and civil liberties more systematically than criminal abuse.

@@

%The next chapter focuses on one specific device key escrow proposal. It defines a threat model and analyzes Stefan
%Savage's 2018 proposal, ``Lawful Device Access without Mass Surveillance Risk: A Technical Design Discussion''
%\cite{savage_lawful_2018}.
